Headgroup Structure and Cation Binding in Phosphatidylserine Lipid Bilayers

### 2019-09-10 Manuscript/manuscriptPS.tex
Clarification concerning systems with Ca2+ counterions.

  The additional simulation details are given in the supplementary information.

@@

  lipid/counter-ions  & force field & {[}CaCl$_{2}${]}\,(M)  & N$_{{\rm l}}$  & N$_{{\rm w}}$  & \footnote{Number of Ca$^{2+}$ cations in the system}N$_{{\rm c}}$  & T (K)  & t$_{{\rm sim}}$(ns)  & t$_{{\rm anal}}$ (ns)  & files\tabularnewline

@@

POPC:POPS (5:1)  & CHARMM36 \cite{klauda10,venable13} / NBfix1 \cite{kim16}  & 0.26  & 250:50 & 11190  & 53  & 298  & 200  & 180  & \citenum{POPC5POPS1withCaClCHARMM} \tabularnewline
POPC:POPS (5:1)  & CHARMM36 \cite{klauda10,venable13} / NBfix1 \cite{kim16}  & 1.06  & 250:50 & 11174  & 214  & 298  & 200  & 180  & \citenum{POPC5POPS1with1MCaClCHARMM} \tabularnewline

@@

POPC:POPS (5:1)  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 0.50  & 120:24 & 5760  & 52  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline
POPC:POPS (5:1)  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 1.00  & 120:24  & 5760  & 104  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline
POPC:POPS (5:1)  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 2.00  & 120:24  & 5760  & 208  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline
POPC:POPS (5:1)  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 3.00  & 120:24 & 5760  & 311  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline
POPC:POPS (5:1)  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 4.00  & 120:24  & 5760  & 415  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline

@@

POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04}  & 0.102\footnote{\label{noteBerger}Calculation of concetration complicated by the

@@

POPC:POPS (4:1)  & Berger \cite{tieleman99,mukhopadhyay04}  & 0.715\textsuperscript{\ref{noteBerger}}  & 104:26 & 4306  & 72  & 310  & 300  & 100  & \citenum{POPCpopsBERGERwith715mMCa} \tabularnewline

@@

\begin{figure*}[tbp]
  \centering
  \includegraphics[width=\textwidth]{../Figs/CHANGESwithCaClPS.eps}
  \caption{\label{changesWITHCaClPS}
   Variation of the POPC (left) and POPS (right) headgroup order parameters in a POPC:POPS (5:1) mixture
    as a function CaCl$_2$ concentration from experiments \cite{roux90} and from different force fields
    at 298~K (except the data for Berger are from a simulation of POPC:POPS (4:1) mixture at 310~K \cite{ollila07a,melcrova16}). 
    The $S_\mathrm{CH}$ values from systems without calcium are set at the zero point of y-axes,
    except for the $\alpha$-carbon order parameters of POPS (bottom right), for which both $S^\alpha_\mathrm{CH}$ are shifted
    such that the lower $S^\alpha_\mathrm{CH}$ is zero without additional ions. This is to correctly illustrate
    the forking with different concentrations of calcium.
    Potassium counterions were used in MacRog simulations and sodium counterions in Lipid14/17 simulations.
    In CHARMM36 and Berger simulations with added calcium, the charge was neutralized with calcium, and monovalent counterions were not present.
  }
\end{figure*}

>>>>

@@

  lipids/counter-ion  & force field & {\footnote{For systems with Ca$^{2+}$ counterions this column gives [Ca$^{2+}$].}[}CaCl$_{2}${]}\,(M)  & N$_{{\rm l}}$  & N$_{{\rm w}}$  & \footnote{Number of Ca$^{2+}$ cations in the system}N$_{{\rm c}}$  & T (K)  & t$_{{\rm sim}}$(ns)  & t$_{{\rm anal}}$ (ns)  & files\tabularnewline

@@

POPC:POPS (5:1)/Ca$^{2+}$  & CHARMM36 \cite{klauda10,venable13} / NBfix1 \cite{kim16}  & 0.26  & 250:50 & 11190  & 53  & 298  & 200  & 180  & \citenum{POPC5POPS1withCaClCHARMM} \tabularnewline
POPC:POPS (5:1)/Ca$^{2+}$  & CHARMM36 \cite{klauda10,venable13} / NBfix1 \cite{kim16}  & 1.06  & 250:50 & 11174  & 214  & 298  & 200  & 180  & \citenum{POPC5POPS1with1MCaClCHARMM} \tabularnewline

@@

POPC:POPS (5:1)/Ca$^{2+}$  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 0.50  & 120:24 & 5760  & 52  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline
POPC:POPS (5:1)/Ca$^{2+}$  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 1.00  & 120:24  & 5760  & 104  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline
POPC:POPS (5:1)/Ca$^{2+}$  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 2.00  & 120:24  & 5760  & 208  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline
POPC:POPS (5:1)/Ca$^{2+}$  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 3.00  & 120:24 & 5760  & 311  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline
POPC:POPS (5:1)/Ca$^{2+}$  & Lipid14/17 \cite{dickson14,gould18} / ff99~\cite{aqvist90}  & 4.00  & 120:24  & 5760  & 415  & 298  & 2$\times$500  & 2$\times$200  & \citenum{POPCpopsLIPID17withCaCl} \tabularnewline

@@

POPC:POPS (4:1)/Ca$^{2+}$  & Berger \cite{tieleman99,mukhopadhyay04}  & 0.102\footnote{\label{noteBerger}Calculation of concentration complicated by the

@@

POPC:POPS (4:1)/Ca$^{2+}$  & Berger \cite{tieleman99,mukhopadhyay04}  & 0.715\textsuperscript{\ref{noteBerger}}  & 104:26 & 4306  & 72  & 310  & 300  & 100  & \citenum{POPCpopsBERGERwith715mMCa} \tabularnewline

@@

\begin{figure*}[tbp]
  \centering
  \includegraphics[width=\textwidth]{../Figs/CHANGESwithCaClPS.eps}
  \caption{\label{changesWITHCaClPS}
   Variation of the POPC (left) and POPS (right) headgroup order parameters in a POPC:POPS (5:1) mixture
    as a function CaCl$_2$ concentration from experiments \cite{roux90} and from different force fields
    at 298~K (except the data for Berger are from a simulation of a POPC:POPS (4:1) mixture at 310~K \cite{ollila07a,melcrova16}). 
    The $S_\mathrm{CH}$ values from systems without calcium are set at the zero point of y-axes,
    except for the $\alpha$-carbon order parameters of POPS (bottom right), for which both $S^\alpha_\mathrm{CH}$ are shifted
    such that the lower $S^\alpha_\mathrm{CH}$ is zero without additional ions. This is to correctly illustrate
    the forking with different concentrations of calcium.
    Counterions were potassium in MacRog simulations, and sodium in Lipid14/17 Dang and CHARMM36 NBfix2 simulations.
    In CHARMM36 NBfix1, Lipid14/17 ff99, and Berger simulations with added calcium the lipid charge was neutralized with calcium, and monovalent counterions were not present;
    for these systems the $x$-axis shows [Ca$^{2+}$].
  }
\end{figure*}
